---
title: "Assignment 1, Task 1"
author: "Elmera Azadpour"
date: "1/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(readxl)
library(here)
library(lubridate)
library(patchwork)
```

### read in Sierra amphibians dataset (provided by The Sierra Lakes Inventory Project)
```{r, echo = TRUE, error = FALSE, warning=FALSE, message=FALSE}
amphibians <- read_excel(here("data","sierra_amphibians.xlsx"))
```
### Graph of total mountain yellow-legged frog (Rana muscosa) count each year across all water bodies
```{r, echo = TRUE, error = FALSE, warning=FALSE, message=FALSE}
amphibian_grouped <- amphibians %>% 
  rename(date = survey_date) %>% 
  mutate(date = ymd(date)) %>% 
  mutate(record_year = year(date)) %>% 
  select(lake_id, amphibian_species, amphibian_life_stage, amphibian_location, record_year) %>% 
  filter(amphibian_species =="RAMU") %>% 
  filter(!amphibian_life_stage =="EggMass") %>% #exclude EggMass level
  group_by(record_year, amphibian_life_stage) %>% 
  count()

plot1 <- ggplot(amphibian_grouped, 
                aes(x = record_year, y = n)) + 
  geom_line(aes(col = amphibian_life_stage)) +
  labs(y = "Count", x = "Year") +
  scale_color_discrete(name = "Amphibian Life Stage",
                       labels = c("Adult", "Sub Adult", "Tadpole")) +
  theme_minimal() +
  theme(legend.position= "top")
```

### Column graph containing total counts of combined adult and subadult endangered mountain yellow-legged frogs (Rana muscosa) observed in the 5 lakes with the greatest total observed counts
```{r, echo = TRUE, error = FALSE, warning=FALSE, message=FALSE}
amphibians_greatest_counts <- amphibians %>% 
  mutate(lake = c("Lake")) %>% 
  unite(col = lake_id,
        c(lake, lake_id), 
        sep = " " ) %>% 
  filter(!amphibian_life_stage =="Tadpole") %>% 
  filter(amphibian_species =="RAMU") %>% 
  select(lake_id, amphibian_life_stage) %>% 
  group_by(lake_id) %>% 
  summarize(count = n()) %>% 
  slice_max(count, n = 5)

plot2 <- ggplot(amphibians_greatest_counts, 
       aes(x =  fct_reorder(lake_id, count), y = count)) +
  geom_col(fill = "dark green", col = "black") +
labs(y = "Total counts adult and subadult Rana muscosa frogs ", 
     x = "Lake ID") +
  coord_flip() +
  theme_minimal()
```

### Patchworked graph
```{r, fig.cap = "**Figure 1**. (A) total mountain yellow-legged frog (*Rana muscosa*) count each year across all bodies of water, by life stage excluding ‘EggMass’ level (B) total counts of combined adult and subadult *Rana muscosa* observed in five lakes with the greatest total observed counts", fig.width= 10, echo = TRUE, error = FALSE, warning=FALSE, message=FALSE}
plot1 / plot2 + plot_annotation(tag_levels = 'A')
```
